---
title: "MATH 392 Problem Set 6"
author: "Sam D. Olson"
output: 
  pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, fig.align = "center", 
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(knitr)
library(here)
```


## Exercise from the book

**9.1 #1**

***Q:*** Let X have the exponential distribution with parameter $\beta$. Suppose that we wish to test the hypotheses $H_0 : \beta \geq 1$ versus $H_1 : \beta < 1$. Consider the test procedure $\delta$ that rejects $H_0$ if $X \geq 1$. 

**(A)**

Determine the power function of the test.  

Let $\delta$ demote a test, specifically: 

$\delta$ :{Reject $H_0$ when $X \geq 1$}

Then for $\beta >0$ we note the power function as follows, $\pi \left( \beta \mid \delta \right) = Pr \left(X \geq 1 \mid \beta \right) = e^{-\beta}$

**(B)**

Compute the size of the test. 

By the definition of size, we have: 

$size(\delta) = sup_{\beta \geq 1} \pi \left( \beta \mid \delta \right)$

Note the power function from **(A)**, $\pi \left( \beta \mid \delta \right)  = e^{-\beta}$, such that $\pi \left( \beta \mid \delta \right)$ is a decreasing function of $\beta$. Thus, the power function takes a maximum at $\beta = 1$, and we may say: 

The size of the test is $\pi \left( \beta = 1 \mid \delta \right) = e^{-1}$.


**9.1 #2**

***Q:*** Suppose that $X_1, ..., X_n$ form a random sample from the uniform distribution on the interval $[0, \theta]$, and that the following hypotheses are to be tested: 

$H_{0} : \theta \geq 2$

$H_{1} : \theta < 2$

Let $Y_n = max \left(X_1, ..., X_n \right)$ and consider a test procedure such that the critical region contains all the outcomes for which $Y_n \leq 1.5$. 

**(A)**

Determine the power function of the test.  

Note, for $0 < y < \theta$, then $Pr \left(Y_n \leq y\right) = \left( \frac{y}{\theta}\right) ^n$.

Additionally note, if $y \geq \theta$, then $Pr \left(Y_n \leq y\right) = 1$.

Thus, for the stated condition, if $\theta \leq 1.5$, then $\pi(\theta) = Pr(Y_n \leq 1.5) = 1$. 

And if $\theta > 1.5$, then $\pi(\theta) = Pr(Y_n \leq 1.5) = \left( \frac{1.5}{\theta} \right) ^n$

**(B)**

Determine the size of the test. 

The size of the test, $size(\delta) = \alpha = sup_{\theta \geq 2} \pi \left(\theta \right) = sup_{\theta \geq 2} \left(\frac{1.5}{\theta} \right)^n = \left(\frac{1.5}{2} \right)^n = \left(\frac{3}{4} \right)^n$.

**9.1 #14** 

Plus: plot power functions in R

***Q:*** Let $X_1, ..., X_n$ be i.i.d. with exponential distribution with parameter $\theta$. Suppose that we wish to test the hypotheses: 

$H_{0} : \theta \geq \theta_0$

$H_{1} : \theta < \theta_0$

Let $X = \sum \limits_{i=1}^{n} {X_i}$. Let $\delta_c$ be the test that rejects $H_0$ if $X \geq c$.

**(A)**

Show that $\pi \left( \theta \mid \delta_c \right)$ is a decreasing function of $\theta$ 

**(B)**

Find c in order to make $\delta_c$ have size $\alpha_0$. 

**(C)**

Let $\theta_0 = 2$, $n=1$, and $\alpha_0 = 0.1$. Find the precise form of the test $\delta_c$ and sketch its power function. 

```{r echo=TRUE, eval=TRUE, fig.heigh=3}
theta <- 250
n <- (1:50)

mle_mse <- rep(NA, 50) 
mom_mse <- rep(NA, 50)
mle_adj_mse <- rep(NA, 50)
for (i in 1:50){ 
  mle_mse[i] <- ((theta ^ 2) * 2 )/ ((n[i] + 1) * (n[i] + 2))
  mom_mse[i] <- (theta ^ 2) / (3 * n[i])
  mle_adj_mse[i] <- (theta ^ 2) / ((n[i] * (n[i] +2)))
}

df <- data.frame(n,mle_mse,mom_mse, mle_adj_mse)

ggplot(df, aes(n)) +                  
  geom_line(aes(y=mle_mse), colour="red") +  
  geom_line(aes(y=mom_mse), colour="green") +
  geom_line(aes(y=mle_adj_mse), colour="blue") +
  geom_text(x=3.5, y=500, label="MLE adj.") +
  geom_text(x=20, y=2500, label="MOM") +
  geom_text(x=8, y=3000, label="MLE") + 
  ggtitle("MSE of MLE, MOM, and adjusted MLE Estimators") +
  labs(y="MSE", x = "n (observations)")
```

**9.2 #2**

***Q:*** Consider two p.d.f.'s $f_{0} \left(x\right)$ and $f_{1} \left(x\right)$ that are defined as follows: 

$$\begin{aligned}
f_{0} \left(x\right)) = \begin{Bmatrix} 
{1} \ for\ 0 \leq x \leq 1 \\
0 \ \ otherwise
\end{Bmatrix}
\end{aligned}$$

and 

$$\begin{aligned}
f_{1} \left(x\right) = \begin{Bmatrix} 
{2x} \ for\ 0 \leq x \leq 1 \\
0 \ \ otherwise
\end{Bmatrix}
\end{aligned}$$

Suppose that a single observation X is taken from a distribution for which the p.d.f. $f(x)$ is either $f_{0} \left(x\right)$ or $f_{1} \left(x\right)$, and the following simple hypotheses are to be tested: 

$H_{0} : f(x) =  f_{0} \left(x\right)$

$H_{1} : f(x) =  f_{0} \left(x\right)$

**(A)**

Describe a test procedure $\delta$ for which the value of $\alpha \left( \delta \right) + 2\beta \left(\delta \right)$ is a minimum. 

Using Thm. 9.2.1 with $a-1, b=2$, we have: 

$\delta$ :{Do not reject $H_0$ if $\frac{f_1(x)}{f_0(x)} < \frac{1}{2}$}

Note, as  $\frac{f_1(x)}{f_0(x)} = 2x \rightarrow 2x < \frac{1}{2}$

Thus we may say the procedure is: Do not reject $H_0$ if $x < \frac{1}{4}$, and likewise to reject $H_0$ if $x > \frac{1}{4}$. 

**(B)**

Determine the minimum value of $\alpha \left( \delta \right) + 2\beta \left(\delta \right)$ attained by that procedure.  

Note the following formulations: 

$\alpha (\delta) = Pr( Reject \space H_0 \mid f_0) = \int \limits_{\frac{1}{4}}^{1} {f_0(x) dx } = \frac{3}{4}$

$\beta (\delta) = Pr( Fail \space to \space  Reject \space  H_0 \mid f_1) = \int \limits_{0}^{\frac{1}{4}} {2x dx } = \frac{1}{16}$

Therefore, we have: 

$\alpha \left( \delta \right) + 2\beta \left(\delta \right) = \frac{3}{4} + 2\frac{1}{16} = \frac{7}{8}$

**9.2 #3**

***Q:*** Consider again the conditions of Exercise 2 (9.2.2), but suppose now that it is desired to find a test procedure for which the value of $3\alpha \left( \delta \right) + \beta \left(\delta \right)$ is a minimum. 

**(A)**

Determine the procedure.  

Using Thm. 9.2.1. with $a=3, b=1$, we have:

$\delta$ :{Do not reject $H_0$ if $\frac{f_1(x)}{f_0(x)} < 3 \rightarrow 2x < 3$}

Since all values of X lie in the interval (0, 1), the optimal procedure is to not reject $H_0$ for every possible observed value (as $max(X)=1 \rightarrow 2 \cdot max(X)=2 < 3$).

**(B)**

Determine the minimum value of $3\alpha \left( \delta \right) + \beta \left(\delta \right)$ attained by the procedure. 

Since we never reject $H_0$ under the stated conditions, we have:

$\alpha (\delta) = 0, \beta(\delta)=1$ and we thus have:

$3\alpha (\delta) + \beta(\delta) = 1$

**9.2 #10**

***Q:*** Suppose that $X_1, ..., X_n$ form a random sample from the Poisson distribution with unknown mean $\lambda$. Let $\lambda_0$ and $\lambda_1$ be specified values such that $\lambda_1 > \lambda_0 > 0$, and suppose that it is desired to test the following simple hypotheses:  

$H_{0} : \lambda = \lambda_0$

$H_{1} : \lambda = \lambda_1$

**(A)**

Show that the value of $\alpha \left( \delta \right) + \beta \left(\delta \right)$ is minimized by a test procedure which rejects $H_0$ when $\bar{X}_n > c$.

**(B)**

Find the value of c. 

**(C)**

For $\lambda_0 = \frac{1}{4}$, $\lambda_1 = \frac{1}{2}$, and $n=20$, determine the minimum value of $\alpha \left( \delta \right) + \beta \left(\delta \right)$ that can be attained. 

**9.3 #1**

***Q:*** Suppose that $X_1, ..., X_n$ form a random sample from the Poisson distribution with unknown mean $\lambda$ $\left( \lambda > 0 \right)$. Show that the joint p.f. of $X_1, ..., X_n$ has a monotone likelihood ratio in the statistic $\sum \limits_{i=1}^{n} {X_i}$.

**9.3 #2**

***Q:*** Suppose that $X_1, ..., X_n$ form a random sample from the normal distribution with known mean $\mu$ and unknown variance $\sigma^2$ $\left( \sigma^2 > 0 \right)$. Show that the joint p.d.f. of $X_1, ..., X_n$ has a monotone likelihood ratio in the statistic $\sum \limits_{i=1}^{n} {\left(X_i - \mu\right)^2}$. 

**9.3 #13**

***Q:*** Suppose that four observations are taken at random from the normal distribution with unknown mean $\mu$ and known variance 1. Suppose also that the following hypotheses are to be tested: 

$H_{0} : \mu \geq 10$

$H_{1} : \mu < 10$

**(A)**

Determine a UMP test at the level of significance $\alpha_{0} = 0.1$

**(B)**

Determine the power of this test when $\mu = 9$. 

**(C)**

Determine the probability of not rejecting $H_0$ if $\mu = 11$. 